Differential invariant signatures for planar Lie group transformations of images

### 2020-08-21 paper/newpaper.tex
Draft of tensor contraction section

h

@@

/subsection{Transformation notation}
Most of the time we will work on grey-scale transformations of two-dimensional images. In this setting, we define the transformation group action by a coordinate transformation $\varphi \colon \mathbb{R}^2 \to \mathbb{R}^2$
\begin{equation}\label{eq:transformation}
  \mathbf{x} \mapsto \bar{\mathbf{x}} = \varphi(\mathbf{x})
\end{equation}
The transformed image is then defined to be
\begin{equation}\label{eq:fbarequalsf}
  \bar{f}(\bar{\mathbf{x}}) = (f\circ\varphi^{-1})(\bar{\mathbf{x}}) = f(\mathbf{x}).
\end{equation}
Derivatives of $\bar{f}$ and $f$ can be related by the chain rule, for example the first order derivatives are given by:
\begin{equation}\label{eq:prolongation}
  \begin{aligned}
    f_x &= \bar{f}_{\bar{x}} \bar{x}_x + \bar{f}_{\bar{y}}\bar{y}_x \\
    f_y &= \bar{f}_{\bar{x}} \bar{x}_y + \bar{f}_{\bar{y}}\bar{y}_y \\
  \end{aligned}
\end{equation}
where $\bar{x}_x = \frac{\partial \varphi_1}{\partial x}(x, y)$, et cetera. It is often convenient to work with $\varphi^{-1}$ instead, which we will do in \S\ref{sec:movingframes}.

====
Let $G$ be a (finite or infinite-dimensional) local planar transformation
group.  A {\em $k$-colour image} is a triple $(f,\Omega,k)$ where
$\Omega\subset\R^2$, $1\le k\in\mathbb{Z}$, and $f\colon\Omega\to\R^k$. The
function $f$ will be taken to be as smooth as necessary.  A transformation
$\varphi\in G$ acts on images by $\varphi\cdot (f,\Omega,k) =
(f\circ\varphi^{-1},\varphi(\Omega),k)$.  (As we work locally, we will
usually omit the domain $\Omega$; if the number of colours $k$ is not
given, we assume $k=1$.)  A differential invariant is an invariant of the
prolonged action of $G$ on the derivatives of $f$, i.e., a function $I$ of
the values of $f$ and its derivatives at a point such that $I(f)(x,y) =
I(\varphi\cdot f)(\varphi(x,y))$. An {\em $m$-dimensional signature set} is
the image of $m$ invariants $\mathcal{I}=(I_1,I_2,\dots,I_m)$, i.e. the set
$\mathcal{I}(f)(\Omega)\subset \mathbb{R}^m$.  A signature set is complete
for $f$ if it determines $f$ up to transformations from $G$, i.e. if
$\mathcal{I}(\tilde f)(\tilde\Omega) = \mathcal{I}(f)(\Omega)$ implies
there exists a transformation $\varphi\in G$ such that $\tilde f =
\varphi\cdot f$ and $\tilde \Omega = \varphi(\Omega)$.

@@

\section{The Invariant Tensor Theorem}

The new contraction mapping computation.

E(2), Sim(2)

Since this will be the first example, put the details of the image here. And remove one of the repeats below.

3D and 4D signatures.

>>>>

@@

\subsection{Transformation notation}
Most of the time we will work on grey-scale transformations of two-dimensional images. In this setting, we define the transformation group action by a coordinate transformation $\varphi \colon \mathbb{R}^2 \to \mathbb{R}^2$
\begin{equation}\label{eq:transformation}
  \mathbf{x} \mapsto \bar{\mathbf{x}} = \varphi(\mathbf{x})
\end{equation}
The transformed image is then defined to be
\begin{equation}\label{eq:fbarequalsf}
  \bar{f}(\bar{\mathbf{x}}) = (f\circ\varphi^{-1})(\bar{\mathbf{x}}) = f(\mathbf{x}).
\end{equation}
Derivatives of $\bar{f}$ and $f$ can be related by the chain rule, for example the first order derivatives are given by:
\begin{equation}\label{eq:prolongation}
  \begin{aligned}
    f_x &= \bar{f}_{\bar{x}} \bar{x}_x + \bar{f}_{\bar{y}}\bar{y}_x \\
    f_y &= \bar{f}_{\bar{x}} \bar{x}_y + \bar{f}_{\bar{y}}\bar{y}_y \\
  \end{aligned}
\end{equation}
where $\bar{x}_x = \frac{\partial \varphi_1}{\partial x}(x, y)$, et cetera. It is often convenient to work with $\varphi^{-1}$ instead, which we will do in \S\ref{sec:movingframes}.

\subsection{Stuff from dispaper to tidy}
====
Let $G$ be a (finite or infinite-dimensional) local planar transformation
group.  A {\em $k$-colour image} is a triple $(f,\Omega,k)$ where
$\Omega\subset\R^2$, $1\le k\in\mathbb{Z}$, and $f\colon\Omega\to\R^k$. The
function $f$ will be taken to be as smooth as necessary.  A transformation
$\varphi\in G$ acts on images by $\varphi\cdot (f,\Omega,k) =
(f\circ\varphi^{-1},\varphi(\Omega),k)$.  (As we work locally, we will
usually omit the domain $\Omega$; if the number of colours $k$ is not
given, we assume $k=1$.)  A differential invariant is an invariant of the
prolonged action of $G$ on the derivatives of $f$, i.e., a function $I$ of
the values of $f$ and its derivatives at a point such that $I(f)(x,y) =
I(\varphi\cdot f)(\varphi(x,y))$. An {\em $m$-dimensional signature set} is
the image of $m$ invariants $\mathcal{I}=(I_1,I_2,\dots,I_m)$, i.e. the set
$\mathcal{I}(f)(\Omega)\subset \mathbb{R}^m$.  A signature set is complete
for $f$ if it determines $f$ up to transformations from $G$, i.e. if
$\mathcal{I}(\tilde f)(\tilde\Omega) = \mathcal{I}(f)(\Omega)$ implies
there exists a transformation $\varphi\in G$ such that $\tilde f =
\varphi\cdot f$ and $\tilde \Omega = \varphi(\Omega)$.

@@

\section{Differential invariants of E(2) through tensor contraction}
While we avoid the use of tensors in most of this study, there is a notable exception to be made for the Euclidean transformation group $E(2)$. In this setting we can restrict ourselves to \emph{Cartesian tensors}, where we need make no distinction between covariance and contravariance, they are both the same thing. We use subscript indices exclusively, for this reason.

In this setting, a Euclidean transformation acting on $\mathbb{R}^2$ is given
by an expression of the form:
\begin{equation*}
  \bar{x}_i = a_{ij}x_j + b_j,
\end{equation*}
using the Einstein summation convention. The orthogonality condition is that $a_{ik}a_{jk} = a_{ki}a_{kj} = \delta_{ij}$.

With respect to $\mathbb{R}^2$, a Cartesian tensor of rank $p$, or $p$-tensor is a $2^p$-tuple of real numbers $C_{i_1\ldots i_p}$ whose components transform under a Euclidean transformation as
\begin{equation*}
  \bar{C}_{i_1\ldots i_p} = a_{i_1j_1}\cdots a_{i_pj_p} C_{j_1\ldots j_p}
\end{equation*}
and a $0$-tensor is a scalar, or invariant.

The main observation is that in this setting, the $n^\text{th}$ order partial derivative operator $\partial^n / \partial_{x_{i_1}} \cdots \partial_{x_{i_n}}$ formally transforms as an $n$-tensor. \todo[inline]{Cite Florack et al, 1993} Because the product of a $p$-tensor and a $q$-tensor gives a $(p+q)$-tensor and contraction of a $p$-tensor yields a $(p-2)$-tensor, a complete contraction of a tensor of even rank formed as a product of terms of the form $\partial f / \partial_{i_1} \cdots \partial_{i_p}$ will give a Euclidean invariant.

For example, up to second derivatives, we have the following Euclidean invariants, expressed as tensor contractions:
\begin{align*}
  f &= f \\
  f_if_i &= f_x^2 + f_y^2 \\
  f_{ii} &= f_{xx}^2 + f_{yy}^2 \\
  f_{ij} f_{ij} &= f_{xx}^2 + 2f_{xy}^2 + f_{yy}^2 \\
  f_i f_j f_{ij} &= f_x^2 f_{xx} + 2f_x f_y f_{xy} + f_y^2 f_{yy}
\end{align*}

While we can construct any number of invariants this way, there is no guarantee of functional independence of the constructed invariants. We will address this issue in Section~\ref{sec:movingframes} when we construct invariants of $E(2)$ and other groups using the method of moving frames.

\todo[inline]{Expand to Sim(2)}
\todo[inline]{Introduce image that we are going to use in subsequenct examples}
\todo[inline]{Introduce signatures - 3D and 4D perhaps?}
